---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
activity<-read.csv("activity.csv")
activity$date<-as.Date(as.character(activity$date))
```
## What is mean total number of steps taken per day?
```{r}
library(dplyr)
library(ggplot2)
```
### calculate the total number of steps taken per day
```{r}
activity<-group_by(activity, date)
totalsteps<-summarise(activity, sum(steps))
```
### make a histogram of the total number of steps taken each day
```{r}
names(totalsteps)[2]<-'totalsteps'
ggplot(data = totalsteps, mapping = aes(x=totalsteps))+
    geom_histogram()
```

### calculate the mean and median of the total number of steps taken per day
```{r}
mean(totalsteps$totalsteps, na.rm = TRUE)
median(totalsteps$totalsteps, na.rm = TRUE)
```

## What is the average daily activity pattern?
### calculate average number of steps taken daily
```{r}
activity<-group_by(activity, interval)
averagesteps<-summarise(activity, mean(steps, na.rm = TRUE))
```
### make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
names(averagesteps)[2]<-'average'
ggplot(data = averagesteps, mapping = aes(x=interval, y=average))+
    geom_line()
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
averagesteps[which.max(averagesteps$average),][1]
```
So interval 835 contains the maximum number of steps.

## Imputing missing values
### calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(activity$steps))
```
### Create a new dataset that is equal to the original dataset but with the missing data filled in
My strategy is to fill in missing values with mean for 5-minute interval which is already calculated above. Because the mean value is the best to reflect typical activities for a given period.
```{r}
activity_new<-activity
activity_new$steps<-ifelse(is.na(activity_new$steps), averagesteps$average, activity_new$steps)
```
### make a histogram of the total number of steps taken each day
```{r}
activity_new<-group_by(activity_new, date)
totalsteps_new<-summarise(activity_new, sum(steps))
names(totalsteps_new)[2]<-'totalsteps'
ggplot(data = totalsteps_new, mapping = aes(x=totalsteps))+
    geom_histogram()
```

### Calculate and report the mean and median total number of steps taken per day
```{r}
mean(totalsteps_new$totalsteps)
median(totalsteps_new$totalsteps)
```
So the mean values are the same. The median value is slightly bigger than that from the fisrt part.
Imputing missing data does not have much impact on the estimates.

## Are there differences in activity patterns between weekdays and weekends?
### create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r}
weekend<-c('Saturday','Sunday')
activity_new$weekday<-ifelse(weekdays(activity_new$date) %in% weekend, 'weekend', 'weekday')
```
### make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days
```{r}
activity_new<-group_by(activity_new, weekday, interval)
averagesteps_new<-summarise(activity_new, mean(steps))
names(averagesteps_new)[3]<-'average'
ggplot(data = averagesteps_new, mapping = aes(x=interval, y=average))+
    geom_line()+
    facet_grid(weekday~.)
```